---
title: 'Monkey ID: Keith (M01)'
subtitle: 'Periodicity analysis'
author: 'Peter Kamerman'
date: "Lasted knitted: `r format(Sys.Date(), '%d %b %Y')`"
---

```{r setup, include = FALSE}
# Load packages
library(magrittr)
library(tidyverse)
library(ggfortify)
library(ggridges)
library(lubridate)
library(multitaper)
library(skimr)
library(zoo)

# Make figure directory
if(!dir.exists('figures/M01-periodicity')) {
    dir.create('figures/M01-periodicity')
}

# knitr setup
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      fig.retina = 2,
                      fig.align = 'center',
                      fig.path = 'figures/M01-periodicity/')
```

----

This script includes analyses on whether there is evidence of circadian variation in body temperature in vervet monkeys. The analysis is for one animal [**Keith (M01)**] and includes chronograms, actograms, periodgrams, and autocorrelation.

----

## Import and summarise data

### Import

```{r import_data}
# Body temperature data
data <- read_rds('data-cleaned/monkey/M01-cleaned.rds')
```

### Summarise

```{r quick_look, results = 'asis'}
kable(skim(data))
```

----

## Process data

```{r process_data}
#-- Generate separate day, month, and year columns --#
data %<>%
    mutate(day = day(date),
           month = month(date, 
                         label = TRUE),
           year = as.character(year(date)))

#-- Smooth body temp data --#
## Rolling smooth over 60 minutes (alignment = centered)
data %<>%
    mutate(temp_smooth = round(rollapply(data = body_temp,
                                   FUN = mean,
                                   width = 12,
                                   partial = TRUE,
                                   align = 'center'), 2))

#-- Define animal ID --#
animal <- str_to_title(data$ID[[1]])
```

----

## Daily summary line plots

```{r daily_summary, fig.height = 8, fig.width = 7}
#-- Body temp --#
data_summary <- data %>%
    # Group and summarise
    group_by(date) %>%
    summarise(mean = mean(body_temp, na.rm = TRUE),
              min = min(body_temp, na.rm = TRUE),
              max = max(body_temp, na.rm = TRUE)) %>%
    ungroup() %>%
    # Add dummy time column and rejoin with data
    mutate(time = '12:00:00') %>%
    left_join(data) %>%
    # Clean-up
    select(date, mean, min, max) %>% 
    # Add body temp label
    mutate(type = 'Body temperature') %>% 
    # Format date column
    mutate(date = ymd(date))

#-- Plot --#
ggplot(data = data_summary) +
    aes(x = date) +
    geom_ribbon(aes(ymax = max, ymin = min), fill = 'wheat2') +
    geom_line(aes(y = mean)) +
    geom_line(aes(y = max), colour = 'wheat3') +
    geom_line(aes(y = min), colour = 'wheat3') +
    labs(title = str_glue('{animal}: Daily mean and range in body and black globe temperature'),
         subtitle = 'Black line: daily mean \nShaded area: daily range',
         x = 'Date',
         y = expression(Body~temperature~(degree*C))) +
    scale_x_date(date_breaks = '2 months',
                 date_labels = '%b %Y') +
    theme_bw(base_size = 12) +
    theme(axis.text.x = element_text(angle = 45,
                                     hjust = 1),
          panel.grid.minor = element_blank())
```
----

## Chronograms

As a preliminary assessment of whether there may be circadian osscilation in body temperature, we plotted chronograms of raw and smoothed _(rolling 60-minute average)_ body temperature data for each month in a calendar year. 

### Year facetted by month

```{r chronogram_ymd, fig.height = 10, fig.width = 8}
#-- Nest data by year --#
chronograms <- data %>% 
    group_by(year) %>% 
    nest()

#-- Plot chronograms by month within each year --#
chronograms %<>%
    # Add dummy date and date-time for plotting
    mutate(data = map(.x = data,
                      ~ .x %>% 
                          mutate(dummy_date_time = ymd_hms(paste0('2000-01-', day, ' ', time)),
                                 dummy_date = ymd_hms(paste0('2000-01-', day, ' 00:00:00'))))) %>% 
    # Plot
    mutate(chronograms = pmap(.l = list(data, year, animal),
                              ~ ggplot(data = ..1) +
                                  aes(x = dummy_date_time,
                                      y = body_temp) +
                                  geom_vline(aes(xintercept = dummy_date), 
                                             colour = '#2678B2') +
                                  geom_point(shape = 21,
                                             stroke = 0.2,
                                             fill = '#FFFFFF') +
                                  geom_line(aes(y = temp_smooth),
                                            colour = '#FD7F28') +
                                  labs(title = str_glue('{..3}: {..2} monthly chronogram'), 
                                       subtitle = 'Open circles: 5-minute temperature data \nOrange line: Smoothed temperature data (60-minute rolling mean, center aligned) \nBlue vertical lines: Midnight of each day',
                                       x = 'Day of month',
                                       y = expression(Body~temperature~(degree~C))) +
                                  scale_x_datetime(date_breaks = '2 days',
                                                   expand = c(0, 0),
                                                   date_labels = '%d') +
                                  scale_y_continuous(limits = c(34, 42)) +
                                  facet_grid(month ~ .,
                                             drop = FALSE) +
                                  theme_bw(base_size = 12) +
                                  theme(panel.grid.minor = element_blank())))

# Print plots
walk(chronograms$chronograms, 
     ~print(.x))
```

----

## Actograms

We then refined the visual inspection of the data for evidence of a circadian rhythm of body temperature by constructing monthly actograms that plotted difference in body temperature _(averaged across 48 30-minute bins per day)_ from the daily mean body temperature for each day of a given month.  

```{r actograms, fig.height = 10, fig.width = 8}
#-- Nest data by year --#
actograms <- data %>% 
    mutate(month = month(date_time, 
                         label = TRUE,
                         abbr = FALSE),
           month = as.character(month)) %>% 
    group_by(year, month) %>% 
    nest() %>% 
    mutate(month = factor(month, 
                          levels = c('January', 'February', 'March', 'April',
                                     'May', 'June', 'July', 'August', 'September',
                                     'October', 'November', 'December',
                                     ordered = TRUE)))

#-- Plot chronograms by month within each year --#
actograms %<>%
    # Prepare for plotting
    mutate(month = as.character(month)) %>% 
    mutate(data_binned = map(.x = data,
                             ~ .x %>% 
                                 # Create 30-minute bins
                                 ## Extract minutes
                                 mutate(time_min = minute(date_time),
                                        time_hour = hour(date_time)) %>% 
                                 ## Re-level minutes to 30 minute intervals
                                 mutate(time_30 = ifelse(time_min > 0 & time_min < 35,
                                                         yes = '00:00',
                                                         no = '30:00'),
                                 time_30 = str_glue('{time_hour}:{time_30}')) %>% 
                                 # Bin by 30-min interval and calculate 
                                 # average temperature per bin 
                                 group_by(day, time_30) %>% 
                                 summarise(temp_binned = mean(body_temp, na.rm = TRUE)) %>% 
                                 # Calculate delta bin and mean daily temperature
                                 group_by(day) %>% 
                                 mutate(temp_delta = temp_binned - mean(temp_binned)) %>% 
                                 ungroup() %>% 
                                 # Make time_30 a date-time object
                                 mutate(time_30 = str_glue('2000-01-01 {time_30}'),
                                        time_30 = ymd_hms(time_30)) %>% 
                                 # Generate a colour aesthestic for the 
                                 # direction of temp_delta
                                 mutate(col_ours = ifelse(temp_delta < 0,
                                                          yes = 'Negative',
                                                          no = 'Positive')) %>% 
                                 # Convert day to an ordered factor
                                 mutate(day = factor(day,
                                                     levels = 1:31,
                                                     ordered = TRUE)))) %>% 
    # Plot data
    mutate(actograms = pmap(.l = list(data_binned, month, year, animal),
                           ~ ggplot() +
                               geom_col(data = ..1,
                                        aes(x = time_30,
                                            y = temp_delta,
                                            fill = col_ours,
                                            colour = col_ours)) +
                               labs(title = str_glue('{..4}: {..2} {..3} daily actograms'), 
                                    subtitle = 'Recordings averaged across 30-minute bins | Unshaded area: Sunrise to sunset',
                                    x = 'Time of day',
                                    y = expression(Delta~body~temperature~from~daily~average~(degree~C))) +
                               scale_x_datetime(date_breaks = '180 mins',
                                                date_labels = '%H:%M',
                                                expand = c(0, 0)) +
                               scale_y_continuous(breaks = 0) +
                               scale_fill_manual(name = expression(Direction~of~Delta),
                                                 values = c('#FD7F28', '#2678B2')) +
                               scale_colour_manual(name = expression(Direction~of~Delta),
                                                   values = c('#FD7F28', '#2678B2')) +
                               facet_grid(day ~ .,
                                          drop = FALSE) +
                               theme_bw(base_size = 12) +
                               theme(axis.text.y = element_blank(),
                                     panel.grid = element_blank(),
                                     panel.spacing.y = unit(0, 'lines'))))

# Print plots
walk(actograms$actograms, 
     ~print(.x))
```

### Summary

The figure below is the monthly average of 5-minute difference in body temperature from mean daily temperature recordings for each year. 

**Please note that the sequence of months is June 2014 (orange) through to June 2015 (blue).**

```{r actogram_summary, fig.height = 10, fig.width = 8}
#-- Prepare summary data for plotting --#
actogram_summary <- actograms %>%
    # Select the required columns
    select(year, month, data) %>% 
    # Unnest the data
    unnest() %>% 
    # Group by year/month/day
    group_by(year, month, day) %>% 
    # Calculate mean daily temperature 
    mutate(mean_temp = mean(body_temp)) %>% 
    # Calculate difference between 5-min temperatures and mean daily temperature
    mutate(delta_temp = body_temp - mean_temp) %>% 
    # Re-group by year/month/time of day
    group_by(year, month, time) %>% 
    # Calculate mean at each time point across each month for each year
    summarise(summary_temp = mean(delta_temp)) %>% 
    ungroup() %>% 
    # Format year/month/time columns
    mutate(month = fct_relevel(factor(month),
                               'January', 'February', 'March', 'April',
                               'May', 'June', 'July', 'August', 'September',
                               'October', 'November', 'December'),
           time = hms(time))

#-- Plot --# 
ggplot() +
    geom_area(data = actogram_summary,
              aes(x = time,
                  y = summary_temp,
                  fill = year),
              colour = '#000000',
              size = 0.2,
              position = 'identity', 
              alpha = 0.6) +
    scale_x_time(breaks = seq(from = 0, 
                              to = 86400, 
                              by = 21600), # Time measured in seconds
                 labels = c('00:00', '06:00', '12:00', 
                            '18:00', '24:00')) + 
    scale_fill_manual(name = 'Year',
                      values = c('#FD7F28', '#2678B2', '#339F34')) +
    scale_colour_manual(name = 'Year',
                        values = c('#FD7F28', '#2678B2', '#339F34')) +
    labs(title = str_glue('{animal}: Average monthly difference in body temperature from\nmean daily temperature for each year'),
         subtitle = 'Shaded area: Sunrise to sunset for the first (yellow) and last (tan) day of the month',
         x = 'Time of day',
         y = expression(Delta~body~temperature~from~daily~mean~(degree*C))) +
    facet_grid(month ~ .,
               drop = FALSE) +
    theme_bw(base_size = 12) +
    theme(panel.grid.minor = element_blank(),
          panel.spacing.y = unit(0, 'lines'))
```

----

## Timing of daily minima and maxima

These analyses use hourly averages of 5-minute body temperature recordings. Median hourly temperature led to too many daily times with the same value and therefore multiple daily minima and maxima. So, we used the trimean, which produces a wider range of hourly temperature values, reducing the number of duplicate values for daily minima and maxima, while also avoiding skewing caused by acute excursions in body temperature that may occur if the arithmetic mean was used.

### Density plots

**Please note that the sequence of months is June 2014 (orange) through to June 2015 (blue).**

```{r timing_density, fig.height = 8, fig.width = 8}
#-- Define custom trimean function --#
trimean <- function(x, na = TRUE) {
    Q25 <- quantile(x, na.rm = na, probs = 0.25)[[1]]
    Q50 <- quantile(x, na.rm = na, probs = 0.50)[[1]]
    Q75 <- quantile(x, na.rm = na, probs = 0.75)[[1]]
    TM <- (2 * Q50) + Q25 + Q75
    TM <- TM / 4
    return(TM)
}

#-- Calculate hourly trimean temp (slow)--#
data_hour <- data %>%
    # Round minutes to the hour (floor)
    mutate(date_hour = floor_date(date_time, unit = 'hour')) %>%
    # Group by data_hour and calculate trimean
    group_by(date_hour) %>%
    summarise(tri_mean = trimean(body_temp)) %>%
    ungroup()

#-- Find time of the minimum/maximum hourly trimean temp for each day --#
# Minima
data_min <- data_hour %>%
    separate(col = date_hour,
             into = c('date', 'hour'),
             sep = ' ',
             remove = FALSE) %>%
    group_by(date) %>%
    filter(tri_mean == min(tri_mean,
                           na.rm = TRUE)) %>%
    ungroup()

# Maxima
data_max <- data_hour %>%
    separate(col = date_hour,
             into = c('date', 'hour'),
             sep = ' ',
             remove = FALSE) %>%
    group_by(date) %>%
    filter(tri_mean == max(tri_mean,
                           na.rm = TRUE)) %>%
    ungroup()

#-- Are there days with multiple minima/maxima? --#
# Minima
multi_min <- data_min %>%
    group_by(date) %>%
    # Get days with multiple minima
    summarise(count = n()) %>%
    filter(count > 1) %>%
    ungroup() %>%
    # Make them pretty
    unite(col = 'days',
          date, count,
          sep = ' (n = ') %>%
    mutate(days = paste0(days, ')')) %>%
    .$days %>%
    paste(., collapse = ', ') %>%
    str_replace_all(pattern = '(.{39})\\s', # stuff in () is a group
                    replacement = '\\1\n') # \1 replaces the group

# Maxima
multi_max <- data_max %>%
    group_by(date) %>%
    # Get days with multiple minima
    summarise(count = n()) %>%
    filter(count > 1) %>%
    ungroup() %>%
    # Make them pretty
    unite(col = 'days',
          date, count,
          sep = ' (n = ') %>%
    mutate(days = paste0(days, ')')) %>%
    .$days %>%
    paste(., collapse = ', ') %>%
    str_replace_all(pattern = '(.{39})\\s', # stuff in () is a group
                    replacement = '\\1\n') # \1 replaces the group

#-- Prepare data_min/max for plotting --#
# Minima
data_min %<>%
    # Extract year, month and hour of day
    mutate(year = year(date),
           year = as.character(year)) %>% 
    mutate(month = month(date, 
                         label = TRUE, 
                         abbr = FALSE),
           month = fct_rev(month)) %>%
    mutate(hr = hour(date_hour)) 

# Maxima
data_max %<>%
    # Extract year, month and hour of day
    mutate(year = year(date),
           year = as.character(year)) %>% 
    mutate(month = month(date, 
                         label = TRUE, 
                         abbr = FALSE),
           month = fct_rev(month)) %>%
    mutate(hr = hour(date_hour)) 

#-- Generate plot objects --#
# Minima
ggplot(data = data_min) +
    aes(x = hr,
        y = month,
        fill = year,
        colour = year) +
    geom_density_ridges2(scale = 0.95, 
                         alpha = 0.6) +
    scale_x_continuous(breaks = c(0, 3, 6, 9, 12, 15, 18, 21),
                       labels = c('00:00', '03:00', '06:00',
                                  '09:00', '12:00', '15:00',
                                  '18:00', '21:00'),
                       limits = c(0, 23)) +
    scale_fill_manual(name = 'Year',
                      values = c('#FD7F28', '#2678B2', '#339F34')) +
    scale_colour_manual(name = 'Year',
                      values = c('#FD7F28', '#2678B2', '#339F34')) +
    labs(title = str_glue('{animal}: Monthly density plot of time of body temperature minima'),
         subtitle = '(Minima extracted from trimean hourly body temperature data)',
         caption = str_glue('Days with multiple minima:\n {multi_min}'),
         x = 'Hour of the day',
         y = 'Month') +
    theme_bw(base_size = 12) +
    theme(panel.grid.minor = element_blank(),
          plot.caption = element_text(size = 10,
                                      hjust = 0))

# Maxima
ggplot(data = data_max) +
    aes(x = hr,
        y = month,
        fill = year,
        colour = year) +
    geom_density_ridges2(scale = 0.95,
                         alpha = 0.6) +
    scale_x_continuous(breaks = c(0, 3, 6, 9, 12, 15, 18, 21),
                       labels = c('00:00', '03:00', '06:00',
                                  '09:00', '12:00', '15:00',
                                  '18:00', '21:00'),
                       limits = c(0, 23)) +
    scale_fill_manual(name = 'Year',
                      values = c('#FD7F28', '#2678B2', '#339F34')) +
    scale_colour_manual(name = 'Year',
                      values = c('#FD7F28', '#2678B2', '#339F34')) +
    labs(title = str_glue('{animal}: Monthly density plot of time of body temperature maxima'),
         subtitle = '(Maxima extracted from trimean hourly body temperature data)',
         caption = str_glue('Days with multiple minima:\n {multi_max}'),
         x = 'Hour of the day',
         y = 'Month') +
    theme_bw(base_size = 12) +
    theme(panel.grid.minor = element_blank(),
          plot.caption = element_text(size = 10,
                                      hjust = 0))
```

### Heatmap

Unlike the density plots, the days with multiple minima or maxima were excluded from the heatmap analysis.

**Please note that the sequence of months is June 2014 through to June 2015.**

```{r timing_heatmap, fig.height = 9, fig.width = 9}
#-- Label maxima and minima data --#
# Minima
data_min %<>%
    mutate(value = 'minimum')

# Maxima
data_max %<>%
    mutate(value = 'maximum')

#-- Filter out days with multiple max/min values --#
# The spread function (used later) does not cope with duplicates

# Fix multi_max/min for use as a filter
## Minima
filter_min <- str_split(string = multi_min,
                        pattern = '\\s\\(.{5}\\)[,]\\s')[[1]] %>%
    str_replace(pattern = '\\s\\(.{5}\\)',
                replacement = '')

## Maxima
filter_max <- str_split(string = multi_max,
                        pattern = '\\s\\(.{5}\\)[,]\\s')[[1]] %>%
    str_replace(pattern = '\\s\\(.{5}\\)',
                replacement = '')

#-- Filter dataframes --#
# Minima
min_filtered <- data_min %>% 
    filter(!date %in% filter_min)

# Maxima
max_filtered <- data_max %>% 
    filter(!date %in% filter_max)

#-- Generate vector of removed days --#
multi_minmax <- c(filter_min, filter_max) %>%
    # Filter for strings on length 10
    .[str_detect(., '.{10}')] %>%
    # Extract unique values
    unique(.) %>%
    # Sort by date
    sort(.) %>%
    # Collapse into a single string
    paste(., collapse = ', ') %>%
    # Add line-breaks
    str_replace_all(pattern = '(.{39})\\s',
                    replacement = '\\1\n')

#-- Join min and max dataframes and organise data --#
data_minmax <- min_filtered %>% 
    bind_rows(max_filtered) %>%
    # clean-up columns
    select(date, hour, month, value) %>%
    # Spread value column
    spread(key = value,
           value = hour) %>%
    # Retain complete cases only
    filter(complete.cases(.))

#-- Calculate time in minutes between min and max --#
data_minmax %<>%
    mutate(interval = interval(paste(date, minimum),
                               paste(date, maximum)),
           length_min = interval / dminutes(1))

#-- Plot --#
data_minmax %>%
    select(-interval) %>%
    mutate(maximum = hour(hms(maximum)),
           minimum = hour(hms(minimum)),
           month = fct_rev(month)) %>%
    arrange(date) %>%
    ggplot(data = .) +
    aes(x = minimum,
        y = maximum) +
    geom_hex(bins = 12) +
    geom_abline(intercept = 0,
                slope = 1,
                colour = '#999999') +
    scale_y_continuous(limits = c(0, 24),
                       expand = c(0, 0),
                       breaks = seq(0, 24, 4)) +
    scale_x_continuous(limits = c(0, 24),
                       expand = c(0, 0),
                       breaks = seq(0, 24, 4)) +
    scale_fill_viridis_c(name = 'Days') +
    labs(title = str_glue('{animal}: Hexagonal heatmap of time of daily minima and maxima'),
         subtitle = 'Black diagonal line: line of identity',
         caption = str_glue('Removed days with multiple maxima or minima: \n{multi_minmax}'),
         x = 'Time of day of minimum (hours)',
         y = 'Time of day of maximum (hours)') +
    facet_wrap(~ month, 
               ncol = 3,
               drop = FALSE) +
    theme_bw(base_size = 12) +
    theme(panel.grid.minor = element_blank(),
          plot.caption = element_text(size = 10,
                                      hjust = 0))
```

### Time between daily minimum and maximum

Unlike the density plots, the days with multiple minima or maxima were excluded from the heatmap analysis.

**Please note that the sequence of months is June 2014 (orange) through to 2014 (blue).**

```{r timing_time, fig.height = 9, fig.width = 9}
#-- Generate plot --#
# Interval calculated in previous code block
data_minmax %>% 
    # Extract year
    mutate(year = year(date),
           year = as.character(year)) %>% 
    ggplot(data = .) +
    aes(x = length_min,
        y = month,
        colour = year,
        fill = year) +
    geom_density_ridges2(scale = 0.95,
                         alpha = 0.6) +
    geom_vline(xintercept = 0,
               colour = 'red',
               size = 1) +
    labs(title = str_glue('{animal}: Density plot of the time difference between\ndaily minimum and maximum body temprature'),
         caption = str_glue('Removed days with multiple maxima or minima:\n {multi_minmax}'),
         x = 'Time difference (min - max; minutes)',
         y = 'Season') +
    scale_fill_manual(name = 'Year',
                      values = c('#FD7F28', '#2678B2', '#339F34')) +
    scale_colour_manual(name = 'Year',
                      values = c('#FD7F28', '#2678B2', '#339F34')) +
    scale_x_continuous(limits = c(-1800, 1800),
                       breaks = seq(-1800, 1800, 600)) +
    theme_bw(base_size = 12) +
    theme(panel.grid.minor = element_blank(),
          plot.caption = element_text(size = 10,
                                      hjust = 0))
```

----

## Periodogram

Periodograms provide an estimate of the spectral density of a signal, with the assumption being that if there is a circadian rhythm, there will be a peak in density around a frequency of 24 hours. To limit the impact that changes in photoperiod across the year may have on the phase of the rhythm, we calculated monthly periodograms rather than assessing the complete data record. We used the multitaper method of signalling processing, with seven tapers and a time-bandwidth parameter of four. 

```{r periodograms}
#-- Nest data by year and month--#
periodgrams <- data %>% 
    mutate(month = month(date_time, 
                         label = TRUE,
                         abbr = FALSE),
           month = as.character(month)) %>% 
    group_by(year, month) %>% 
    nest() %>% 
    mutate(month = factor(month, 
                          levels = c('January', 'February', 'March', 'April',
                                     'May', 'June', 'July', 'August', 'September',
                                     'October', 'November', 'December',
                                     ordered = TRUE)))

#-- Generate periodogram data for each month --#
periodgrams %<>% 
    # Perform spectral analysis of the time series 
    # using the multitaper method (using default settings: nw = 4, k = 7)
    mutate(periodogram = map(.x = data,
                             ~ spec.mtm(ts(.x$body_temp), 
                                        nw = 4, k = 7, 
                                        dtUnits = 'day', 
                                        deltat = 1/288,
                                        jackknife = TRUE,
                                        plot = FALSE,
                                        log = 'yes'))) %>% 
    # Add freq, spec, and jacknife 95% CIs into a dataframe
    mutate(periodogram_df = map(.x = periodogram,
                                ~ tibble(frequency = .x$freq, 
                                         spectrum = .x$spec,
                                         upper_CI = .x$mtm$jk$upperCI,
                                         lower_CI = .x$mtm$jk$lowerCI)))
    
#-- Generate periodgram and F-value plots for each month --#
## Days per month
days <- map(.x = periodgrams$data, 
            ~ round(nrow(.x) / 288))
## Plot
periodgrams %<>%
    # Periodogram plots
    mutate(periodogram_plot = pmap(.l = list(periodogram_df, 
                                             as.character(month), 
                                             year, animal, days),
                                   ~ if(..5 < 16) {
                                       ggplot(data = ..1) +
                                       aes(x = frequency, 
                                           y = spectrum) +
                                       geom_blank() + 
                                       labs(title = str_glue('{..4}: {..2} {..3} periodogram'), 
                                            subtitle = str_glue('No plot: series length < 16 days'),
                                            x = 'Period (days)',
                                            y = 'Spectral density') +
                                       scale_x_continuous(limits = c(0, 5),
                                                          expand = c(0, 0),
                                                          breaks = c(0.01, 0.5, 1, 
                                                                     2, 3, 4, 5),
                                                          labels = as.character(round(c(1/0.01, 1/0.5, 1/1, 1/2, 1/3, 1/4, 1/5), 2))) +
                                       theme_bw(base_size = 12) +
                                       theme(panel.grid.minor = element_blank())
                                   } else {
                                       ggplot(data = ..1) +
                                       aes(x = frequency, 
                                           y = spectrum) +
                                       geom_ribbon(aes(ymin = lower_CI, 
                                                       ymax = upper_CI),
                                                   fill = 'wheat2',
                                                   colour = 'wheat3') +
                                       geom_line() + 
                                       labs(title = str_glue('{..4}: {..2} {..3} periodogram'), 
                                            subtitle = str_glue('Series length (days): {..5} \nShaded area: 95% jackknife confidence interval \nMultitaper method: time-bandwidth (nw) = 4, tapers (k) = 7'),
                                            x = 'Period (days)',
                                            y = 'Spectral density') +
                                       scale_x_continuous(limits = c(0, 5),
                                                          expand = c(0, 0),
                                                          breaks = c(0.01, 0.5, 1, 
                                                                     2, 3, 4, 5),
                                                          labels = as.character(round(c(1/0.01, 1/0.5, 1/1, 1/2, 1/3, 1/4, 1/5), 2))) +
                                       theme_bw(base_size = 12) +
                                       theme(panel.grid.minor = element_blank())}))
    
# Print plots
walk(periodgrams$periodogram_plot,
     ~print(.x))
```

----

## Autocorrelation

Like the periodogram analysis, to limit the impact that changes in photoperiod across the year may have on the phase of the rhythm, we calculated autocorrelation for monthly data rather than assessing the complete data record. 

In addition, we reduced the size and smoothed the data before performing the autocorrelation by averaging the 5-minute body temperature data across 48 30-minute duration bins per day. The autocorrelation was performed with a maximum lag of 480 30-minute bins (10 days). 

```{r autocorrelation}
#-- Nest data by year and month--#
autocorr <- data %>% 
    mutate(month = month(date_time, 
                         label = TRUE,
                         abbr = FALSE),
           month = as.character(month)) %>% 
    group_by(year, month) %>% 
    nest() %>% 
    mutate(month = factor(month, 
                          levels = c('January', 'February', 'March', 'April',
                                     'May', 'June', 'July', 'August', 'September',
                                     'October', 'November', 'December',
                                     ordered = TRUE)))

# Process data
autocorr %<>% 
    # Prepare for plotting
    mutate(month = as.character(month)) %>% 
    # Create 30-minute binned body temperature data
    mutate(temp_binned = map(.x = data,
                             ~ .x %>% 
                                 # Extract minutes
                                 mutate(time_min = minute(date_time),
                                        time_hour = hour(date_time)) %>% 
                                 # Re-level minutes to 30 minute intervals
                                 mutate(time_30 = ifelse(time_min > 0 & time_min < 35,
                                                         yes = '00:00',
                                                         no = '30:00'),
                                 time_30 = str_glue('{time_hour}:{time_30}')) %>% 
                                 # Bin by 30-min interval and calculate 
                                 # average temperature per bin 
                                 group_by(day, time_30) %>% 
                                 summarise(temp_binned = mean(body_temp, na.rm = TRUE)) %>% 
                                 ungroup() %>%  
                                 # Extract body_binned
                                 .$temp_binned)) %>% 
    # Generate acf object (autocorrelation lag.max = 10 days (n = 480))
    mutate(ACF = map(.x = temp_binned,
                     ~ acf(.x, 
                           lag.max = 480,
                           plot = FALSE))) %>%   
    # Is the length of temp_binned less than 10 days (480 bins)
    mutate(data_length = map(.x = temp_binned,
                             ~ ifelse(length(.x) < 480,
                                      yes = 'TRUE',
                                      no = 'FALSE'))) %>%   
    # Plot data
    mutate(acf_plots = pmap(.l = list(ACF, as.character(month), year, animal, data_length),
                           ~ if(..5 == 'FALSE') {
                               autoplot(..1,
                                        conf.int.fill = '#2678B2') +
                               labs(title = str_glue('{..4}: {..2} {..3} autocorrelation of body temperature'),
                                    subtitle = 'Temperature data averaged across 30-minute bins \nMaximum lag: 10 days (480 bins) \nBlue ribbon: 95% confidence interval of a white noise process',
                                    x = 'Days',
                                    y = 'Autocorrelation coefficient') +
                               scale_y_continuous(breaks = round(seq(from = -0.6,
                                                                     to = 1,
                                                                     by = 0.2), 1),
                                                  limits = c(-0.6, 1.1),
                                                  expand = c(0, 0)) +
                               scale_x_continuous(breaks = seq(from = 0, 
                                                               to = 480, 
                                                               by = 48),
                                                  labels = 0:10,
                                                  limits = c(-5, 485),
                                                  expand = c(0,0)) +
                               theme_bw(base_size = 12) +
                               theme(panel.grid.minor = element_blank())
                               
                               } else {
                                   df <- data.frame(acf = ..1$acf[1:nrow(..1$acf), , 1],
                                                    lag = ..1$lag[1:nrow(..1$lag), , 1])
                                   
                                   ggplot(data = df) +
                                       aes(x = lag,
                                           y = acf) +
                                       geom_blank() +
                                       labs(title = str_glue('{..4}: {..2} {..3} autocorrelation of body temperature'),
                                            subtitle = 'No Plot: Number of days < maximum lag period (10 days)',
                                            x = 'Days',
                                            y = 'Autocorrelation coefficient') +
                                       scale_y_continuous(breaks = round(seq(from = -0.6,
                                                                             to = 1,
                                                                             by = 0.2), 1),
                                                          limits = c(-0.6, 1.1),
                                                          expand = c(0, 0)) +
                                       scale_x_continuous(breaks = seq(from = 0, 
                                                                       to = 480, 
                                                                       by = 48),
                                                          labels = 0:10,
                                                          limits = c(-5, 485),
                                                          expand = c(0,0)) +
                                       theme_bw(base_size = 12) +
                                       theme(panel.grid.minor = element_blank())}))

# Print plots
walk(autocorr$acf_plots, 
     ~print(.x))
```

----

## Session information

```{r session_info}
sessionInfo()
```
